Combining living and fossil taxa into phylogenies: the missing data issue

### 2015-07-11 Manuscript/MPESubmission/Revision/response.tex

\item{\textcolor{blue}{\textbf{line 273:} 'Bayesian Inference'}}

We changed the title as suggested.

\item{\textcolor{blue}{\textbf{line 286:} Why was the true tree used as a starting tree? Earlier the authors argued that the true tree is almost never known, so using it here detracts from the generalizability of these results. If the only justification is 'to speed up the Bayesian estimation process' as stated later, then it seems like this may bias one of the main results, which is that Bayesian analyses outperformed max. likelihood, in which no starting tree was used. I understand the computational limitations, but the authors should present some analysis with complete and missing data in which no starting tree is given compared to analyses in which a starting tree is given to evaluate whether using a starting tree significantly affects the resulting trees, perhaps in an Appendix. Or at least include some perturbation to the starting tree to avoid overly influencing the resulting topology. It would seem to me that this is illustrated in figure 5, where-in with the Robinson-Foulds metric, the distance to the best tree in Bayesian analysis seems to have a lower bound at ~0.7, even under the worst cases of missing data. This may be the worst the analysis can do given that it started out with the true tree. }}

We added an extra analysis on a subset of five trees with the three missing data parameters at 0\%, 10\%, 25\%, 50\% and 75\% for 20 chains. We compared the topology of these trees in Bayesian inference when using the true tree as a starting tree or using a random tree (default in MrBayes) as a starting tree. We see no difference in topologies between both the Bayesian inferences starting with the true tree or a random tree. We made these data available in the supplementary materials Appendix A @@@ and we also added a precision in the manuscript:

blabalbalbal. line @@@ 
%To do (MAJOR)

### 2015-07-11 Manuscript/MPESubmission/Revision/OnlineAppendices-LaTeXSuppFiles/AppendixA.tex
response update

>>>>

\subsection{Effect of the starting tree in Bayesian inference}
We used the ``true'' tree as a starting tree for the two MCMC chains in the Bayesian inference for decreasing the computing time.
To assess if this had an effect on the topology of the ``best'' tree, we ran a sub-sample of trees using a different random tree for the two MCMC chains \citep[default MrBayes option;][]{Ronquist2012mrbayes}.
We tested this effect on five trees with the five levels of missing data (i.e. first tree: $M_L$=0\%, $M_F$=0\% and $M_C$=0\%; second tree: $M_L$=10\%, $M_F$=10\% and $M_C$=10\%, etc.) on the first 20 simulation chains.
We then compared the trees inferred using the ``true'' tree as a starting tree to the trees using a random tree as a starting tree using the normalised Robinson-Foulds and Triplets metrics.
Additionally, we compared the trees inferred using a random tree as a starting tree to the ``best'' tree (inferred using the ``true'' tree as a starting tree) and to the ``true'' tree to see whether our results varied if we didn't used the ``true'' tree as a starting tree.

\begin{table}[!ht]
\caption{Normalised topological metric between trees using the ``true'' or a random tree as a starting tree for the MCMC chains.}
\centering
\begin{tabular}{|l|c|c|c|c|c|c|c|}
  \hline
 Tree &  Metric & Min. & 1st Qu. & Median & Mean & 3rd Qu. & Max. \\ 
  \hline
    $M_L$=0\%; $M_F$=0\%; $M_C$=0\%          & $RF$ & 0.00 & 0.00 & 0.10 & 0.20 & 0.32 & 1.00 \\ 
                                             & $Tr$ & 0.34 & 0.49 & 0.61 & 0.62 & 0.75 & 1.00 \\ 
    $M_L$=10\%; $M_F$=10\%; $M_C$=10\%       & $RF$ & 0.03 & 0.54 & 0.69 & 0.64 & 0.77 & 0.98 \\ 
                                             & $Tr$ & 0.08 & 0.57 & 0.65 & 0.64 & 0.73 & 0.82 \\ 
    $M_L$=25\%; $M_F$=25\%; $M_C$=25\%       & $RF$ & 0.02 & 0.74 & 0.80 & 0.79 & 0.89 & 0.98 \\ 
                                             & $Tr$ & 0.21 & 0.67 & 0.73 & 0.72 & 0.77 & 0.84 \\ 
    $M_L$=50\%; $M_F$=50\%; $M_C$=50\%       & $RF$ & 0.00 & 0.00 & \textbf{0.00} & \textbf{0.01} & 0.01 & 0.04 \\ 
                                             & $Tr$ & 0.08 & 0.38 & 0.59 & 0.57 & 0.73 & 0.84 \\ 
    $M_L$=75\%; $M_F$=75\%; $M_C$=75\%       & $RF$ & 0.00 & 0.00 & \textbf{0.01} & \textbf{0.02} & 0.04 & 0.11 \\ 
                                             & $Tr$ & 0.21 & 0.36 & 0.56 & 0.55 & 0.74 & 0.87 \\ 
   \hline
\end{tabular}
\label{Tab_Results-Difference_methods}
\end{table}

%Incl table + incl figure
